---
title: "12: JSAbacusFramework.io"
---

```{r}
#| message: false
library(mistlecode)
library(jsonlite)
dt <- readLines("input.json")
```

## Part 1

Surprisingly easy. Just had a small regex typo with `?` instead of `+`.

```{r}
dt |>
  str_extract_all("(-?[0-9]+)") |>
  unlist() |>
  as.numeric() |>
  sum(na.rm = TRUE)
```

## Part 2

What a nightmare! I got the answer wrong 16 times and tried so many things. String replacement, string reduction, listing, unlisting, whatever I could think of. I eventually remembered rrapply and had a wacky solution that didn't work at all. I took a break for a few days and really read the documentation on `how = "recurse"`. Then it was just a matter of realizing that `.xname` in the documentation is not the same as `names(x)` and I was good to go!

```{r}
dt |>
  parse_json() |>
  list() |>
  rrapply::rrapply(
    \(o) { any(o == "red") & any(names(o) %in% letters) },
    \(x) { return(list("a" = "red")) }, 
    how = "recurse",
    classes = "list") |>
  unlist() |>
  as.numeric() |>
  suppressWarnings() |>
  sum(na.rm = TRUE)
```
