---
title: "2022-14: Regolith Reservoir"
format: html
---

This doesn't actually look too terrible.

```{r}
#| message: false
library(mistlecode)
```

```{r}
dt <- 
  readLines("input.txt") |>
  str_split(" -> ")
```

This was a standalone in Part 1, but I moved it to a function for Part 2.

```{r}
createCave <- function(maxRow, maxCol) {
  cave <- matrix(" ", maxRow, maxCol)
  
  for (j in dt) {
    for (i in 2:length(j)) {
      x <-
        str_split(j[i - 1], ",", simplify = TRUE)[1, ] |>
        as.numeric()
      y <-
        str_split(j[i], ",", simplify = TRUE)[1, ] |>
        as.numeric()
      
      cave[x[2]:y[2], x[1]:y[1]] <- "█"
    }
  }
  return(cave)
}
```

## Part 1

First step is always making the matrix way bigger than it needs to be and
filling it. Because everything's a straight line, I can just match the current
pair to the next pair and draw a line, essentially. Super easy with matrices. I
want to check the bottom row for sand and if there isn't sand, drop a new one
from `(1,500)` (hooray for R being y,x and non-zero indexed!). Instead of
iterating through the whole drop, I can just get the index of the last row in
column 500 where there isn't anything. This saves some computation. If I'm at
the bottom, write a piece of sand. From there it's just iterating over
everything just like the instructions say. Lastly, get the amount of sand in the
cave then subtract one to account for the last piece of sand that trips the
loop.

```{r}
maxCol <-
  dt |>
  str_extract_all("[0-9]{1,3},") |>
  suppressWarnings() |>
  unlist() |>
  str_remove_all(",") |>
  as.numeric() |>
  max()
cave <- createCave(maxCol, maxCol)
while(!any(cave[nrow(cave),] == "o")) {
  col <- 500
  row <- which(cave[,col] != " ")[1] - 1
  breakFlag <- FALSE
  repeat {
    if (row + 1 > nrow(cave)) { breakFlag <- TRUE }
    else {
      if (cave[row + 1, col] == " ") {
        row <- row + 1
      } else if (col >= 2 & cave[row + 1, col - 1] == " ") {
        row <- row + 1; col <- col - 1
      } else if (col < ncol(cave) & cave[row + 1, col + 1] == " ") {
        row <- row + 1; col <- col + 1
      } else { breakFlag <- TRUE }
    }
    if (breakFlag) {
      cave[row, col] <- "o"
      break
    }
  }
}
which(cave == "o") |>
  length() |>
  sum(-1)
```

## Part 2

Okay. I need to get the highest row, I can't just make a massive matrix. Then
add another row on top of where everything was drawn that will capture the top
of the pyramid. Also, make a new column placeholder that can be added if I need
to expand to the left or right. Because of that, I also need to track where
column 500 is. If I'm going to have a column collision, I need to expand in that
direction. Then, it's business as usual, but now I'm checking to see if that
original start point is full, no matter where it is. Lastly, don't subtract one
because the top grain does count now.

```{r}
maxRow <-
  dt |>
  str_extract_all(",[0-9]{1,3}") |>
  suppressWarnings() |>
  unlist() |>
  str_remove_all(",") |>
  as.numeric() |>
  max()
cave <- createCave(maxRow + 1, maxCol)
cave <- rbind(matrix(" ", 1, ncol(cave)), cave)
newCol <- matrix(" ", nrow(cave), 1)
startCol <- 500
while (cave[1,startCol] != "o") {
  col <- startCol
  row <- which(cave[,col] != " ")[1] - 1
  breakFlag <- FALSE
  repeat {
    if (col == 2) { 
      cave <- cbind(newCol, cave)
      startCol <- startCol + 1
    } else if (col == ncol(cave) - 1) { cave <- cbind(cave, newCol) }
    
    if (row + 1 > nrow(cave)) { 
      breakFlag <- TRUE 
    } else {
      if (cave[row + 1, col] == " ") {
        row <- row + 1
      } else if (col >= 2 & cave[row + 1, col - 1] == " ") {
        row <- row + 1; col <- col - 1
      } else if (col < ncol(cave) & cave[row + 1, col + 1] == " ") {
        row <- row + 1; col <- col + 1
      } else { breakFlag <- TRUE }
    }
    if (breakFlag) {
      cave[row, col] <- "o"
      break
    }
  }
}
which(cave == "o") |>
  length()
```

