---
title: "2022-02: Rock Paper Scissors"
format: html
---

```{r}
library(tidyverse)
library(data.table)
```

```{r}
dt <- fread("input.txt", header = FALSE)
# dt <- readLines("input.txt")
```

This isn't so bad, just rock, paper, scissors.

## Part 1

I initially did `V1 == V2` but that doesn't work with `ABC` and `XYZ`, but a quick fix put me at 549 on the leaderboard so I'm really pleased.

```{r}
dt %>%
  mutate("my_score" = case_when(
    V2 == "X" ~ 1,
    V2 == "Y" ~ 2,
    V2 == "Z" ~ 3
  )) %>%
  mutate("win_score" = case_when(
    (V1 == "A" & V2 == "X") | (V1 == "B" & V2 == "Y") | (V1 == "C" & V2 == "Z") ~ 3,
    V1 == "A" & V2 == "Y" ~ 6,
    V1 == "A" & V2 == "Z" ~ 0,
    V1 == "B" & V2 == "X" ~ 0,
    V1 == "B" & V2 == "Z" ~ 6,
    V1 == "C" & V2 == "X" ~ 6,
    V1 == "C" & V2 == "Y" ~ 0
  )) %>%
  mutate("score" = my_score + win_score) %>%
  pull("score") %>%
  sum()
```

## Part 2

I had this one pretty quick, just a bit of logic to get the `play` column right. Of course, I then had to remember to change all the right `V2` references which took me a minute to figure out when I missed the second mutate.

```{r}
dt %>%
  mutate("play" = case_when(
    V2 == "X" & V1 == "A" ~ "Z",
    V2 == "X" & V1 == "B" ~ "X",
    V2 == "X" & V1 == "C" ~ "Y",
    V2 == "Y" & V1 == "A" ~ "X",
    V2 == "Y" & V1 == "B" ~ "Y",
    V2 == "Y" & V1 == "C" ~ "Z",
    V2 == "Z" & V1 == "A" ~ "Y",
    V2 == "Z" & V1 == "B" ~ "Z",
    V2 == "Z" & V1 == "C" ~ "X",
  )) %>%
  mutate("my_score" = case_when(
    play == "X" ~ 1,
    play == "Y" ~ 2,
    play == "Z" ~ 3
  )) %>%
  mutate("win_score" = case_when(
    (V1 == "A" & play == "X") | (V1 == "B" & play == "Y") | (V1 == "C" & play == "Z") ~ 3,
    V1 == "A" & play == "Y" ~ 6,
    V1 == "A" & play == "Z" ~ 0,
    V1 == "B" & play == "X" ~ 0,
    V1 == "B" & play == "Z" ~ 6,
    V1 == "C" & play == "X" ~ 6,
    V1 == "C" & play == "Y" ~ 0
  )) %>%
  mutate("score" = my_score + win_score) %>%
  pull("score") %>%
  sum()
```

