---
title: "Mull It Over"
format: html
date: "2024-12-03"
---

```{r}
#| message: false
library(mistlecode)

options(scipen = 999)
```

```{r}
dt <- readLines("input.txt")
```

## Part 1

I'm really pleased with this. Doing the `parse() |> eval()` is super super dangerous and should never be done, but it works so I'm happy. I got 343 on the leaderboard which is my best ever.

```{r}
mul <- function(x, y) x * y
dt |>
  stringr::str_extract_all('mul\\(\\d+,\\d+\\)') |>
  unlist() |>
  purrr::map_int(\(x) {
    x |>
      parse(text = _) |>
      eval()
  }) |>
  sum()
```

## Part 2

I didn't realize the example input changed which messed me up a bit. Then I just had to get my regex right and I was good to go. I'm also pleased with my little `if...else if` chain. I got an error when `x` was a `mul()` call, but then realized that I didn't have to explicitly check because I was already checking for `do()` and `don't()` and if it's not either of those, then it must be a `mul()`. Oh, and 502 on the leaderboard for part 2 would've been a best ever if not for part 1!

```{r}
do <- TRUE
dt |>
  stringr::str_extract_all("mul\\(\\d+,\\d+\\)|do\\(\\)|don't\\(\\)") |>
  unlist() |>
  purrr::map_int(\(x) {
    if (x == "don't()") do <<- FALSE 
    else if (x == "do()") do <<- TRUE 
    else if (do) {
      x <-
        x |>
        parse(text = _) |>
        eval()
      return(x)
    }
    return(0)
  }) |>
  sum()
```

