---
title: "20: Infinite Elves and Infinite Houses"
---

```{r}
#| message: false
library(mistlecode)
library(Rcpp)
library(RcppArmadillo)
```

## Part 1

There's something I can do with prime factorization, I just can't figure it out. Ugh. Took me a fat minute but I was able to work with my idea about how if you divide a number `x` by `1:x` then round down, you get the total number of times each elf would have stopped at a house before `x`. Just had to get that set up right and we're good to go. Not the fastest, but it works. Maybe another chance for `Rcpp` if I'm up for it later.

Spoiler alert. I wasn't happy with the plain R version and re-wrote it in `Rcpp`. At this point I've been through so many versions, I think my working `R` solution is lost...

::: {.panel-tabset}

### `Rcpp`

```{r}
cppFunction('
IntegerVector part1(int target, int total_l) {
  IntegerVector m_vec, sum_vec(total_l, 0.0), v;
  IntegerMatrix m_mat;
  int idx = 0;
  while (idx <= total_l) {
    idx++;
    if (idx % 1000 == 0) { std::cout << "\\r" << idx; }
    v = rep_len(IntegerVector::create(idx), total_l / idx);
    m_mat = IntegerMatrix(idx, v.size());
    m_mat(idx - 1, _) = v * 10;
    m_vec = IntegerVector(m_mat);
    m_vec.attr("dim") = R_NilValue;
    m_vec = rep_len(m_vec, total_l);
    sum_vec = sum_vec + m_vec;
    for (int i = 0; i < sum_vec.size(); i++) {
      if (i <= idx & sum_vec[i] >= target) {
        if (idx > 1000) { std::cout.clear(); }
        return IntegerVector::create(i) + 1;
      }
    }
  }
  if (idx > 1000) { std::cout.clear(); }
  return sum_vec;
}
')
part1(36e6, 1e6)
```

### `R`

```{r}
#| eval: false
target <- 36e6 / 10
idx <- 1
s <- 0
d <- seq_len(target)
denominator <- d
repeat {
  denominator <- d[seq_len(idx)]
  s <- sum(denominator[idx %% denominator == 0])
  if (s < target) idx <- idx + 1 else break
}
idx
```

:::

## Part 2

```{r}
cppFunction('
int part2(int target, int total_l) {
  IntegerVector m_vec, sum_vec(total_l, 0.0), v;
  IntegerMatrix m_mat;
  int idx = 0, j = 0;
  while (idx <= total_l) {
    idx++;
    if (idx % 1000 == 0) { std::cout << "\\r" << idx; }
    v = rep_len(IntegerVector::create(idx), total_l / idx);
    v = v * 11;
    m_mat = IntegerMatrix(idx, v.size());
    m_mat(idx - 1, _) = v;
    m_vec = IntegerVector(m_mat);
    //m_vec.attr("dim") = R_NilValue;
    j = 0;
    //std::cout << idx;
    for (int i = total_l - 1; i >= 0; i--) {
      if (m_vec[i] != 0 & j < 50) { 
        j++; 
        m_vec[i] = 0;
      }
    }
    m_vec = rep_len(m_vec, total_l);
    sum_vec = sum_vec + m_vec;
    for (int i = 0; i < sum_vec.size(); i++) {
      if (i <= (idx - 1) & sum_vec[i] >= target) {
        if (idx > 1000) { std::cout.clear(); }
        return(i + 1);
      }
    }
  }
  if (idx > 1000) { std::cout.clear(); }
  return -1;
}
')
part2(120, 100)

cppFunction('
IntegerVector test(int l) {
  IntegerVector x;
  x = seq_len(l);
  x = x * 11;
  for (int i = l - 1; i >= 0; i--) {
    if (x[i] % 2 == 0) {
      x[i] = 0;
    }
  }
  return x;
}
')
test(100)
```

