---
title: "Camp Cleanup"
format: html
date: "2022-12-04"
---

```{r}
library(tidyverse)
library(data.table)
```

```{r}
dt <- fread("input.txt", header = FALSE)
```

This looks bad at first, but it's not.

## Part 1

It took me a minute to figure out how I wanted to work with the data ranges. I started with `mutate` then `lapply` then quickly switched to `mapply`. From there it wrote itself. I then second-guessed my answer and divided by two because I thought I was double reporting the ranges. I cost myself a minute :(

```{r}
mapply(\(x, y) {
  x <- as.numeric(unlist(strsplit(x, "-")))
  y <- as.numeric(unlist(strsplit(y, "-")))
  
  all(x[1]:x[2] %in% y[1]:y[2]) | all(y[1]:y[2] %in% x[1]:x[2])
}, dt$V1, dt$V2) |>
  unlist() |>
  table() %>%
  .["TRUE"]
```

## Part 2

38 seconds for part 2! I just had to change `all` to `any`.

```{r}
mapply(\(x, y) {
  x <- as.numeric(unlist(strsplit(x, "-")))
  y <- as.numeric(unlist(strsplit(y, "-")))
  
  any(x[1]:x[2] %in% y[1]:y[2]) | any(y[1]:y[2] %in% x[1]:x[2])
}, dt$V1, dt$V2) |>
  unlist() |>
  table() %>%
  .["TRUE"]
```

