---
title: "Monkey Math"
format: html
date: "2022-12-21"
---

This...doesn't actually look too bad. Part 2 will probably be a nightmare.

```{r}
#| message: false
library(mistlecode)
#remotes::install_github("rconsortium/OOP-WG")
library(S7)
options(scipen = 999)
```

```{r}
dt <- 
  readLines("input.txt") |>
  str_split(": ")
```

## Part 1

Let's take a chance to use [S7](https://rconsortium.github.io/OOP-WG/)
for the first time! I can build on the bits I learned with S3 in
[2022-11](/2022/11/2022-11.html). The idea to use a vector of monkey
names (`index`) made it really easy to make sure I was pulling info from
the right monkeys, rather than needing to iterate over the entire list.
There's probably a better approach overall with a more recursive
solution, but this seems to work fine for Part 1, although I'm worried
about Part 2.

Note (2023-04-23): I just updated to R 4.3 and the latest version of R7
which is now S7 and I was unable to run the previous version which used
recursion due to a C-stack error. I swapped over to a while loop based
version and that seems to be working fine.

```{r}
monkey <- new_class(
  name = "monkey",
  properties = list(
    name = class_character,
    val = class_any,
    op = class_character
  )
)

new_monkey <- function(x) {
  if (!is.na(suppressWarnings(as.numeric(x[2])))) {
    val <- as.numeric(x[2])
    op <- NA_character_
  } else {
    val <- NA
    op <- x[2]
  }
  monkey(
    name = x[1],
    val = val,
    op = op
  )
}

monkeys <- lapply(dt, new_monkey)
masterMonkeys <- monkeys
index <- sapply(monkeys, \(x) x@name)

monkey_math <- function(monkeys, op) {
  m1 <- monkeys[[which(index == op[1])]]@val
  m2 <- monkeys[[which(index == op[3])]]@val
  
  if (!is.na(m1) & !is.na(m2)) {
    if (op[2] == "+") { m1 + m2 }
    else if (op[2] == "-") { m1 - m2 }
    else if (op[2] == "*") { m1 * m2 }
    else if (op[2] == "/") { m1 / m2 }
  } else { return(NA_integer_) }
}

m <- monkeys
breakFlag <- FALSE
while(!breakFlag) {
  n <- 1
  while (n <= length(monkeys)) {
    if (is.na(monkeys[[n]]@val)) {
      val <- str_split_1(monkeys[[n]]@op, " ")
      monkeys[[n]]@val <- monkey_math(monkeys, val)
    }
    if (monkeys[[n]]@name == "root" & !is.na(monkeys[[n]]@val)) {
      breakFlag <- TRUE
      break
    }
    n <- n + 1
  }
}
monkeys[[n]]@val
```

## Part 2

I tried brute-forcing it... It took a while to fully figure out what I wanted to
do, but the idea behind just making a giant equation then solving for `humn` was
there early on. I started by expanding the input by replacing monkeys with their
value if they had one or operation if they didn't. I then replaced all the
monkeys with their values, just like in Part 1. This left me with a giant
expression where the only variable was `humn`. I just wasn't sure how to
actually solve equations in R so I kept trying to brute-force it, which kept on
not working. I eventually found the `Ryacas` package which made quick work of
everything.

```{r}
monkeys <- lapply(dt, new_monkey)

root_op <- 
  str_extract(monkeys[[which(index == "root")]]@op, "(\\+|\\-|\\*|\\/)")
monkeys[[which(index == "root")]]@op <- 
  gsub(paste0("\\", root_op), "=", monkeys[[which(index == "root")]]@op)

m1 <- str_split_1(monkeys[[which(index == "root")]]@op, " = ")
m2 <- m1[2]
m1 <- m1[1]

replace_monkey <- function(m) {
  w <- which(index == m)
  if (length(w) == 0) { return(m) }
  v <- monkeys[[w]]@val
  if (is.na(v)) { return(paste0("( ", monkeys[[which(index == m)]]@op, " )")) }
  return(paste0("( ", v, " )"))
}

# https://stackoverflow.com/a/14838753
quotemeta <- function(string) { str_replace_all(string, "(\\W)", "\\\\\\1") }

reduce_string <- function(mm, pattern) {
  r <-
    str_extract_all(mm, pattern) |>
    unlist() |>
    sapply(\(r) {
      if (grepl("h", r)) { return(r) }
      else { eval(parse(text = r)) }
    })
  
  for (i in 1:length(r)) {
    mm <- str_replace_all(mm, quotemeta(names(r)[i]), as.character(r[i]))
  }
  return(mm)
}

process_string <- function(m) {
  while (any(grepl("([^(h)][a-z])", m))) {
    m <- 
      sapply(m, \(mm) { replace_monkey(mm) }) |>
      paste0(collapse = " ") |>
      str_split_1(" ")
  }
  m <- paste0(m, collapse = "")
  oldM <- 0
  while (length(m) == 1 && oldM != m) {
    oldM <- m
    m <- reduce_string(m, "\\((\\d+|h)\\)[\\+|\\-|\\*|\\/]\\((\\d+|h)\\)")
  }
  return(oldM)
}

monkeys[[which(index == "humn")]]@val <- "h"
mm1 <- process_string(m1)
mm2 <- process_string(m2)

paste(
    reduce_string(mm1, "\\((\\d+|h)\\)"), 
    "==", 
    reduce_string(mm2, "\\((\\d+|h)\\)")
  ) %>%
  Ryacas::y_fn("Solve", "h") |>
  Ryacas::yac_str() |>
  str_extract("\\d+") |>
  as.numeric()
```
